---
title: "README"
date: "November 5, 2016"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# eCLV

by Glenn Solomon, Shachi Parikh, Jordan McIntyre, Brian Silverman, Tejaswi Namana, Maaz Rehan,
Koba Khitalishivli

# Introduction

Collection of functions that compute Expected Customer Lifetime Value in a subscription (contractual) 
setting based on research by Fader and Hardie.

## Usage

We will use the example data from paper 1 in the references section.

```{r, warning=F}
# Load the functions
source("lib-BG.R")

# Example from paper
# Data
activeCust = c(869,743,653,593,551,517,491)
lostCust = c(131,126,90,60,42,34,26)

# Estimate the maximum likelihood function to get the alpha, beta values
# Alternative function to estimate MLL
# estimateMLL2(active.cust = activeCust,
#             lost.cust = lostCust)$par
alphabeta = estimateMLL(active.cust = activeCust,
            lost.cust = lostCust)$par

alphabeta

# Survival probabilities for 10 periods
survivalBG(alphabeta[1],alphabeta[2],1:10)

# Churn probabilities for 10 periods
churnBG(alphabeta[1],alphabeta[2],1:10)

#  Get the retention rates for 10 periods
retentionRates(alphabeta[1],alphabeta[2],1:10)

# Discounted Expected Lifetime
DEL(alphabeta[1],alphabeta[2],discount = 0.1)

# Discounted Expected Residual Lifetime
DERL(alphabeta[1],alphabeta[2],discount = 0.1, renewals = 4)

# Expected Lifetime Value per customer assuming expected net cash flow is 100$ per period
DEL(alphabeta[1],alphabeta[2],discount = 0.1)*100

# Expected ResidualLifetime Value per customer assuming expected net cash flow is 100$ per period
DERL(alphabeta[1],alphabeta[2],discount = 0.1)*100
```

## References

1. [How to Project Customer Retention](https://www.google.com/url?sa=t&rct=j&q=&esrc=s&source=web&cd=1&cad=rja&uact=8&ved=0ahUKEwjXxq-v4ZLQAhWp64MKHYtSC9UQFggdMAA&url=http%3A%2F%2Fciteseerx.ist.psu.edu%2Fviewdoc%2Fdownload%3Fdoi%3D10.1.1.558.935%26rep%3Drep1%26type%3Dpdf&usg=AFQjCNHiSlM9GzZC_VIbQ2rgtSJ6dtSdwA&sig2=twz37wWrSTSSHAo6Dwj0iA) by Peter Fader and Bruce Hardie (2007)

2. [Fitting the sBG Model to Multi-Cohort Data](http://brucehardie.com/notes/017/) by Peter Fader and Bruce Hardie (2007)
